Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-03-19 WSSSPE2_report/WSSSPE2_review.tex
Section 3.1: First draft.

\usepackage{color}
\usepackage{dcolumn}
\usepackage{float}
\usepackage{graphicx}
\usepackage[latin9]{inputenc}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{subfigure} 
\usepackage{psfrag}
\usepackage{tabularx}
\usepackage[hyphens]{url}
\usepackage{wrapfig}
\usepackage{longtable}

@@

\newcommand{\todo}[1]{{\color{blue}$\blacksquare$~\textsf{[TODO: #1]}}}
\newcommand{\note}[1]{ {\textcolor{blueish}    { ***Note:      #1 }}}
\newcommand{\katznote}[1]{ {\textcolor{magenta}    { ***Dan:      #1 }}}
\newcommand{\gabnote}[1]{ {\textcolor{cyan}    { ***Gabrielle:     #1 }}}
\newcommand{\nchnote}[1]{  {\textcolor{orange}      { ***Neil: #1 }}}
\newcommand{\manishnote}[1]{  {\textcolor{violet}     { ***Manish: #1 }}}
\newcommand{\davidnote}[1]{  {\textcolor{darkgreen}      { ***David: #1 }}}
\newcommand{\colinnote}[1]{ {\textcolor{red}    {***Colin #1 }}}

@@

Thaney started noting the unfortunate fact that many current systems have been designed to create friction hindering us, despite the original intentions. \katznote{prev sentence is confusing}%---social issues hindering us
An example is the National Cancer Institute's caBIG. 
A total of \$350m was spent, including more than \$60m for management.  $70+$ tools were created, but caBIG is still seen
as a failure (Kaitlin didn't really say why? Were the tools used? Did they
immediately bit-rot? Was their development unsustained?)-should ask question
about this.
 Those that had the least investment were the most used, the largest
investments were the least used.
 Is there a caBIG report available? Quotes from Andrea Califano/Joe Gray,
looks like there is:
\url{http://tinyurl.com/maf6dz2}. %http://www.informationweek.com/architecture/report-blasts-problem-plagued-cancer-research-grid/d/d-id/1097068?

@@

Inefficiency: \url{http://www.xkcd.com/1445/}. Need research tools (e.g., software repositories), social capital (e.g.,
incentives) and capacity (e.g., training and mentorship) for efficient reproducible
research.

Our systems need to talk to each other but Q for speaker is that as
systems become less monolithic, paradoxically it may be harder to sustain the
links between them see:
\url{http://tinyurl.com/l76tba2} %http://www.slideshare.net/jameshowison/scientific-software-sustainability-and-ecosystem-complexity
(but does Anon Grizzly have other cites/info for that?).
Note, for example, that creative commons licensed works typically do not 
have dependencies (a book, a photo, an artwork).
Software Discovery Index: \url{http://softwarediscoveryindex.org/report/}.

 Final points:
\begin{itemize}
\item Design with the audience not the individual in mind
\item Design to unlock latent potential of our systems
\item Rethink how we reward researchers and support roles
\item By mindful of jargon/semantics traps
\end{itemize}

Questions:
\begin{itemize}
\item Ted Habermann: if the metadata is minimal less onerous for data provider,
but more onerous for user. How to work with data providers to show that users
need more.
\item Find balance between things that have to be captured and what would be
ideal. Don't want to get to point where like old Harvard Dataverse, three pages
of metadata capture, only first four things ever filled out. Don't want to lose
engagement.
\end{itemize}

>>>>

\usepackage{color}
\usepackage{dcolumn}
\usepackage{float}
\usepackage{graphicx}
\usepackage[latin9]{inputenc}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{subfigure} 
\usepackage{psfrag}
\usepackage{tabularx}
\usepackage[hyphens]{url}
\usepackage{wrapfig}
\usepackage{longtable}
\usepackage{verbatim}

@@

\newcommand{\todo}[1]{{\color{blue}$\blacksquare$~\textsf{[TODO: #1]}}}
\newcommand{\note}[1]{ {\textcolor{blueish}    { ***Note:      #1 }}}
\newcommand{\katznote}[1]{ {\textcolor{magenta}    { ***Dan:      #1 }}}
\newcommand{\gabnote}[1]{ {\textcolor{cyan}    { ***Gabrielle:     #1 }}}
\newcommand{\nchnote}[1]{  {\textcolor{orange}      { ***Neil: #1 }}}
\newcommand{\manishnote}[1]{  {\textcolor{violet}     { ***Manish: #1 }}}
\newcommand{\davidnote}[1]{  {\textcolor{darkgreen}      { ***David: #1 }}}
\newcommand{\colinnote}[1]{ {\textcolor{red}    {***Colin #1 }}}
\newcommand{\choinote}[1]{ {\textcolor{orange}    {***Choi #1 }}}

@@

Thaney started noting the unfortunate fact that many current systems suffer the
unintended consequence of creating friction that hinders users, despite
designers' original purposes to do good. An example is the National Cancer
Institute's caBIG. A total of \$350 million was spent, including more than \$60
million for management. More than $70$ tools were created, but caBIG is still
seen as a failure\footnote{\url{http://tinyurl.com/maf6dz2}}. Those that had the
least investment were the most used; the most invested software were the least
utilized.
\begin{comment}
(Kaitlin didn't really say why? Were the tools used? Did they immediately
bit-rot? Was their development unsustained?)-should ask question about this. Is
there a caBIG report available? Quotes from Andrea Califano/Joe Gray, looks like
there is: \url{tinyurl.com/qdodbo7}.
%http://www.informationweek.com/architecture/report-blasts-problem-plagued-cancer-research-grid/d/d-id/1097068

@@

Inefficiency cartoon:  \url{http://www.xkcd.com/1445/}. 
\end{comment}

Thaney emphasized that for efficient reproducible open research, we would need
research tools (e.g., software repositories), social capital (e.g., incentives),
and capacity (e.g., training and mentorship). Our systems would need to
communicate with each other. A point was made by a member of the auidence that
as systems become less monolithic, it often becomes harder to sustain the links between
them\footnote{See, for example, \url{http://tinyurl.com/l76tba2}.}.
%http://www.slideshare.net/jameshowison/scientific-software-sustainability-and-ecosystem-complexity
%(but does Anon Grizzly have other cites/info for that?).
In contrast, works licensed by Creative Commons typically do not 
have dependencies (a book, a photo, an artwork).

Thaney introduced Mozilla's Software Discovery
Index\footnote{\url{http://softwarediscoveryindex.org/report/}}, which
identifies scientific software by archiving and standardizing metadata for
software and hence help connect both developers and users. Ted Habermann in the
audience commented that if the metadata is minimal, it would be less onerous for
data providers, but more burdensome for users---it could be challenging to keep
a balance between what have to be captured and what would be ideal if we do not
want to lose user engagement as in the case of the old Harvard Dataverse, finding often
only the first four fields of three pages of metadata were filled out.

The speaker concluded her talk urging the audience to design scientific software
with the general community, not an individual, in mind; and to design to unlock
latent potential of our systems. In addition, she encouraged everyone to rethink
how we reward researchers and support roles. Lastly, she cautioned the community
to be mindful of jargon or semantics traps.
